cardinalR: Generating interesting high-dimensional data structures

### 2025-09-16 paper-cardinalR.Rmd
Updated the Quadratic text with "randomly sampling inputs" than Uniform
Updated the cone generation

This function generates $n$ points in $p\text{-}D$, where the last dimension ($p$) represents height along the cone's axis, and the remaining dimensions define a shrinking hyperspherical cross-section as one moves toward the tip. Points along the height axis, $p$, are drawn from an exponential distribution to increase density near the tip. That is, $X_p = h_i \sim \text{Exp}(\lambda = 2/h)$ truncated at $h$. The effective radius at height $h_i$ decreases linearly from base to tip, controlled by a shape ratio parameter $r_i = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}})h_i/h, \quad \text{where } r_{\text{min}} = \text{ratio},\ r_{\text{max}} = 1$.

For each point, a direction is sampled uniformly from a $(p-1)$-dimensional hypersphere using generalized spherical coordinates (angles). The radial coordinates are scaled by $r_i$, which ensures a conical taper. Specifically, when $p = 3$, this scaling results in a classic $3\text{-}D$ cone. However, for $p > 3$, the introduction of additional angular components allows for a smooth extension into higher dimensions, preserving the conical shape while accommodating the complexities of multi-dimensional geometry.

@@

The curve is constructed by drawing uniformly spaced inputs and applying a second-degree polynomial transformation. Let $X_1 \sim U(\text{range}[1], \text{range}[2])$ be the independent variable. A raw polynomial basis of degree 2 is used to form $X_2 = X_1 - X_1^2 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$. This creates a smooth parabolic arc that opens downward, with jitter in the vertical direction to simulate noise. For $p > 2$, Gaussian noise $X_j \sim N(0, 0.1^2)$ is added to embed the $2\text{-}D$ structure into $p\text{-}D$, where $j = 3, \dots, p$.

>>>>

This function generates $n$ points in $p-D$, where the last dimension, $X_p$, represents the height along the cone's axis, and the first $p-1$ dimensions define a shrinking hyperspherical cross-section toward the tip. Heights are sampled from a truncated exponential distribution, $X_p \sim \text{Exp}(\lambda = 2/h)$, capped at the cone height $h$, producing a higher density of points near the tip. At each height $X_p$, the radius of the cross-section decreases linearly from base to tip according to $r = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) X_p / h$, where $r_{\text{min}} = \text{ratio}$ and $r_{\text{max}} = 1$.

For each point, a direction in the first $p-1$ dimensions is sampled uniformly on a $(p-1)$-dimensional hypersphere using generalized spherical coordinates. The radial coordinates are scaled by the height-dependent radius $r$, producing the conical taper. In three dimensions ($p = 3$), this results in a classical 3-D cone, while for $p > 3$, additional angular components allow a smooth extension into higher dimensions, preserving the conical shape within multi-dimensional space.

@@

The curve is constructed by randomly sampling inputs and applying a second-degree polynomial transformation. Let $X_1 \sim U(\text{range}[1], \text{range}[2])$ be the independent variable. A raw polynomial basis of degree 2 is used to form $X_2 = X_1 - X_1^2 + \varepsilon_2$, where $\varepsilon_2 \sim U(0, 0.5)$. This creates a smooth parabolic arc that opens downward, with jitter in the vertical direction to simulate noise. For $p > 2$, Gaussian noise $X_j \sim N(0, 0.1^2)$ is added to embed the $2\text{-}D$ structure into $p\text{-}D$, where $j = 3, \dots, p$.

### 2025-09-16 paper-cardinalR.tex
Updated the Quadratic text with "randomly sampling inputs" than Uniform
Updated the cone generation

This function generates \(n\) points in \(p\text{-}D\), where the last dimension (\(p\)) represents height along the cone's axis, and the remaining dimensions define a shrinking hyperspherical cross-section as one moves toward the tip. Points along the height axis, \(p\), are drawn from an exponential distribution to increase density near the tip. That is, \(X_p = h_i \sim \text{Exp}(\lambda = 2/h)\) truncated at \(h\). The effective radius at height \(h_i\) decreases linearly from base to tip, controlled by a shape ratio parameter \(r_i = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}})h_i/h, \quad \text{where } r_{\text{min}} = \text{ratio},\ r_{\text{max}} = 1\).

For each point, a direction is sampled uniformly from a \((p-1)\)-dimensional hypersphere using generalized spherical coordinates (angles). The radial coordinates are scaled by \(r_i\), which ensures a conical taper. Specifically, when \(p = 3\), this scaling results in a classic \(3\text{-}D\) cone. However, for \(p > 3\), the introduction of additional angular components allows for a smooth extension into higher dimensions, preserving the conical shape while accommodating the complexities of multi-dimensional geometry.

@@

The curve is constructed by drawing uniformly spaced inputs and applying a second-degree polynomial transformation. Let \(X_1 \sim U(\text{range}[1], \text{range}[2])\) be the independent variable. A raw polynomial basis of degree 2 is used to form \(X_2 = X_1 - X_1^2 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\). This creates a smooth parabolic arc that opens downward, with jitter in the vertical direction to simulate noise. For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.1^2)\) is added to embed the \(2\text{-}D\) structure into \(p\text{-}D\), where \(j = 3, \dots, p\).

>>>>

This function generates \(n\) points in \(p-D\), where the last dimension, \(X_p\), represents the height along the cone's axis, and the first \(p-1\) dimensions define a shrinking hyperspherical cross-section toward the tip. Heights are sampled from a truncated exponential distribution, \(X_p \sim \text{Exp}(\lambda = 2/h)\), capped at the cone height \(h\), producing a higher density of points near the tip. At each height \(X_p\), the radius of the cross-section decreases linearly from base to tip according to \(r = r_{\text{min}} + (r_{\text{max}} - r_{\text{min}}) X_p / h\), where \(r_{\text{min}} = \text{ratio}\) and \(r_{\text{max}} = 1\).

For each point, a direction in the first \(p-1\) dimensions is sampled uniformly on a \((p-1)\)-dimensional hypersphere using generalized spherical coordinates. The radial coordinates are scaled by the height-dependent radius \(r\), producing the conical taper. In three dimensions (\(p = 3\)), this results in a classical 3-D cone, while for \(p > 3\), additional angular components allow a smooth extension into higher dimensions, preserving the conical shape within multi-dimensional space.

@@

The curve is constructed by randomly sampling inputs and applying a second-degree polynomial transformation. Let \(X_1 \sim U(\text{range}[1], \text{range}[2])\) be the independent variable. A raw polynomial basis of degree 2 is used to form \(X_2 = X_1 - X_1^2 + \varepsilon_2\), where \(\varepsilon_2 \sim U(0, 0.5)\). This creates a smooth parabolic arc that opens downward, with jitter in the vertical direction to simulate noise. For \(p > 2\), Gaussian noise \(X_j \sim N(0, 0.1^2)\) is added to embed the \(2\text{-}D\) structure into \(p\text{-}D\), where \(j = 3, \dots, p\).
